  \begin{thebibliography}{99}



\bibitem{B}       http://www.woodrow.org/teachers/mi/1993/37burc.html

\bibitem{Bi}      Bittinger, Marvin L., \emph{Calculus and Its Applications}, Addision-Wesley, 2004.

\bibitem{BW}   Barker, William H. and James E.Ward, \emph{The Calculus Companion}, 5th edition, Wiley, 1995.

\bibitem{EP}     Edwards, C. Henry and David E. Penney, \emph{Calculus Early Transcendentals Version}, 6th edition, Prentice Hall, 2003.  (Concept questions are included in every section of the text.)

\bibitem{EP2}	Edwards, C. Henry and David E. Penney, \emph{Calculus Matrix Version}, 6th edition, Prentice Hall, 2003.  (There are a large number of true/false questions on the accompanying CD.)

\bibitem{FWG} Finney, Wier, Giordano, \emph{Calculus}, 10th edition, Addison Wesley Longman, 2001.

\bibitem{MC6}	``Mathematics in College'', Instructional Resource Center and the CUNY Mathematics Discussion Group, City University of New York,  Spring-Summer, 1986.

\bibitem{MC7}	``Mathematics in College'', Instructional Resource Center and the CUNY Mathematics Discussion Group, City University of New York,  Spring-Summer, 1987.

\bibitem{MR}    Meier, John and Thomas Rishel, \emph{Writing in the Teaching and Learning of Mathematics}, Mathematical Association of America, 1998.

\bibitem{R}       Riddle, Douglas F., \emph{Calculus and Analytic Geometry}, 2nd edition, Wadsworth, 1974.

\bibitem{S}       Sterret, Andrew, editor, \emph{Using Writing to Teach Mathematics}, Mathematical Association of America, 1990.

\bibitem{SBS}   Strauss, Bradley, Smith, Calculus, 3rd edition, Prentice Hall, 2002.

\bibitem{SMo}  Smith, David A. and Lawrence C. Moore, \emph{Instructor's Guide with Test Items to accompany Calculus Modeling and Application}, D.C. Heath, 1996.

\bibitem{SM}    Smith, Robert T. and Roland B. Minton, \emph{Calculus}, premiere edition,McGraw Hill, 2000. (Writing prompts are included in every section of this text.)

\bibitem{SM1}  Smith, Robert T. and Roland B. Minton, \emph{Calculus}, McGraw Hill, 2006. (Writing prompts are included in every section of this text.)

\bibitem{SO}    Spiegel, Murray R., \emph{Schaum's Outline of Theory and Problems of Complex Variables}, McGraw-Hill, 1999.

\bibitem{SJ}   Stonewater, Jerry K., ``The Mathematics Writer's Checklist:  The Development of a Preliminary Assessment Tool for Writing in Mathematics'',\emph{ School Science and Mathematics}, Volume 102, Number 7, November 2002.



\end{thebibliography}

\end{document}
