\chapter{}  

 \section{}  \begin{enumerate}  
 
 \item  Consider the function $$f(x) = {x \over {\left| x \right|}}.$$  Calculate $f(x)$ for $x$ = $-2$, $-1.5$, $-1$, $-0.5$, 0, 0.5, 1, 1.5 and 2.  Does $$\mathop {\lim }\limits_{x \to 0} f(x)$$ exist? Discuss.  

\item  Describe the limitations of finding limits by tables or graphs. 

\end{enumerate}
\section{}
\begin{enumerate}


\item  Evaluate $$\mathop {\lim }\limits_{x \to 0} \left[ {x^2  - {{\cos x} \over {1,000,000,000}}} \right]$$ analytically.  Explain why it would be difficult to find this limit with either graphical or numerical methods. \cite{SBS} 

\end{enumerate}
\section{}
\begin{enumerate}



\item  Explain what the Intermediate Value Theorem is, how to use it, and how it is useful. 

\item  In what ways can a function be discontinuous?  Give examples and explain what part of the definition of continuity your examples do not satisfy. 


\item  \label{alaska} Last June I visited my sister in Fairbanks, Alaska.  One Saturday, we hiked up Chena Dome starting at 9 am.  At the top, we made camp and stayed the night.  On Sunday, we started at 9 am and started back down the hill.  Prove that at some time of day on Saturday and Sunday we were at exactly the location on the trail up Chena Dome.\\ 
	$\mbox{}\ \ \ \ \ \ $(Hint:  Draw a sketch of the function of our altitude on Saturday versus time.  On the same graph, draw a sketch of the function of our altitude on Sunday versus time.  A picture is not a proof, but it should help lead you on the right path (pun intended).)\\ 
	$\mbox{}\ \ \ \ \ \ $Note that we cannot say exactly what time of day this "crossing of paths" happened, but that is not the point.  The issue here is simply that it must have happened.  This is called an existence proof.  It is like knowing that your spouse is having an affair.  You may not know who they are having the affair with or when they met up, but you may have proof that they are having an affair, i.e., the affair exists. 

\item  Must there have been some time in your life when your height ($h$) in inches exactly matched your weight ($w$) in pounds?  For most people your age it has.  Prove that at one time they were exactly the same.   Refer to Problem \ref{alaska} for more information. 


\item  What is the definition of continuity?  Explain each part of the definition.  Give an example of 4 functions:  one that is continuous and one that does not fulfill each of the parts of the definition of continuity. 

\end{enumerate}\section{} \begin{enumerate} 

\item  a)  Evaluate the limit $\mathop {\lim }\limits_{x \to 0} {{e^{2x}  - 1} \over {e^x  - 1}}$ and explain each step.  \\
b) Evaluate the limit $\mathop {\lim }\limits_{x \to 0} {{e^{3x}  - 1} \over {e^x  - 1}}$ and explain each step.  \\
c)	Generalize. 

\end{enumerate}


